---
output: pdf_document
---
## Exploring MPG of Select Car Models

```{r, echo=FALSE}
library(datasets)
data(mtcars)
```


### Executive Summary

The purpose of this study is to investigate the relationship between mpg and the
type of transmission as well as the effect of any confounding variables.  We would
specifically like to know whether manual or automatic results in better gas mileage.

Cars with manual transmission will get 2.9 more miles per gallon than cars with automatic transmissions adjusted by weight and quarter mile time.  

### Analysis

As shown by fig1, the mean for an automatic transmission is 17.15 and the mean for a manual transmission is 24.39.  However, this is just based on a simple model. To explore further we'll examine the results of all variables fitted to mpg:

```{r}
allVar <- lm(mpg ~ ., mtcars)
summary(allVar)
```

Using the step function to remove/add each variable iteratively to the regression.   

```{r, results='hide'}
bestVar <- step(allVar, direction = "both")
```

``` {r}
summary(bestVar)
```

The best fit model uses the variables weight (per 1000lbs), qsec (1/4 mile time) and transmission.  The adjusted R-squared of our best fit linear regression is 83.4.  That is to say those three variables explain 83.4% of the variation in mpg.  Fig3 shows the interaction between the 3 variables and mpg, confirming our coefficients.  

-Every 1000 lb increase results in -3.92 mpg

-Every 1 sec increase in qsec results in 1.23 mpg (increasing qsec should mean a decrease in performance)


```{r}
justTransmission <- lm(mpg ~ am, mtcars)
anova(justTransmission, bestVar)
```

Shows the p-value is highly significant, thus rejecting the null hypothesis that wt and qsec don't contribute to mpg.

### Residuals

For the followiing section refer to fig2.:

- Residuals vs Fitted <- points are randomly scattered indicating independence

- Normal QQ <- points fall on line indicating normally distributed

- Scale-Location <- indicates constant variance


```{r}
leverage <- hatvalues(bestVar)
tail(sort(leverage), 3)
```
```{r}
influential <- dfbetas(bestVar)
tail(sort(influential[,3]), 3)
```

- Residuals vs Leverage <- Points mentioned here match cars indicated by the measures of influence. 

### Inference

```{r}
t.test(mpg~am, mtcars)
```

Rejects the null hypothesis that there is no difference in mpg based on transmission.

### Index

```{r, echo=FALSE}
boxplot(mpg ~ am, data = mtcars, xlab = "Transmission: 0 = Automatic, 1 = Manual",
        ylab = "MPG", main = "fig1. MPG by Transmission Type")
```

fig2.
```{r, echo=FALSE}
par(mfrow = c(2, 2))
plot(bestVar)
```

fig3.
```{r, echo=FALSE}
mtcars_vars <- mtcars[ ,c(1,6,7,9)]
pairs(mtcars_vars, panel = panel.smooth)
```

